A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2013-10-25 jacobian_calc_CMAME.tex

\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation with each method.}
For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
central-difference and forward-difference will be shown and explained. The expressions for AD,
central-difference and forward-difference are taken from \cite{ref-Adaggio}, while the expression
for CTSE is original.

When using the forward-difference technique the formula for calculating a tangent-stiffness
matrix is:

\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x)}{\delta}
\end{equation}

Where $K_{ij}$ is the index notation representation of an element of the tangent-stiffness matrix at
row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$
w.r.t the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}.

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_speed.pgf}}
  \caption{Serial test series speed measurements.}
  \label{fig:serial_speed}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_speed.pgf}}
  \caption{Multicore test series speed measurements.}
  \label{fig:multi_speed}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_accuracy.pgf}}
  \caption{Serial test series accuracy measurements.}
  \label{fig:serial_accuracy}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_accuracy.pgf}}
  \caption{Multicore test series accuracy measurements.}
  \label{fig:multi_accuracy}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/serial_speed_rel.pgf}}
  \caption{Serial test series efficiency measurements.}
  \label{fig:serial_efficiency}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.7}{\input{./figs/multi_speed_rel.pgf}}
  \caption{Multicore test series efficiency measurements.}
  \label{fig:multi_efficiency}
\end{figure}

>>>>

\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation.}

When using the forward-difference technique the formula for calculating a tangent-stiffness
matrix is:
%
\begin{equation} 
  K_{ij} = \frac{\partial F_i^{int}(x + h e_j) - F_i^{int}(x)}{h}
\end{equation}
%
Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$ with respect the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}.

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
  \caption{Serial test series speed measurements.}
  \label{fig:serial_speed}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
  \caption{Multicore test series speed measurements.}
  \label{fig:multi_speed}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/serial_accuracy.pgf}}
  \caption{Serial test series accuracy measurements.}
  \label{fig:serial_accuracy}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/multi_accuracy.pgf}}
  \caption{Multicore test series accuracy measurements.}
  \label{fig:multi_accuracy}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/serial_speed_rel.pgf}}
  \caption{Serial test series efficiency measurements.}
  \label{fig:serial_efficiency}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{1.0}{\input{./figs/multi_speed_rel.pgf}}
  \caption{Multicore test series efficiency measurements.}
  \label{fig:multi_efficiency}
\end{figure}
